---
title: 'A Study of Carbon Emissions in the US'
author: "Rogelio Caballero"
date: "November 13, 2016"
output: html_document
---

The following is a commentary to my answers to the Final Project of Exporatory Data Analysis course, Data Science Specilization, Johns Hopkins University on Coursera.

The idea was to make an exploratory analysis of carbon emissions in the US as a function of the variables: county, source of emission and time (1999 - 2008).

## Data and Package imports

The first step was to import all packages that we were going to use through the exercises.

I decided to include *ggplot2* because of it was mandatory for a couple of questions but also for its convenience and intuitive notation.

*dplyr* package was also included because of its elegant notation and efficient  
work with large data frames.

```{r}
library(dplyr)
library(ggplot2)
```

The emissions data was contained in an *.rds* file called *summarySCC_PM25.rds*. Since it is a pretty large file, the use of *dplyr* is justified.
```{r}
p <- "/home/rogelio/Desktop/datasciencecoursera/"
ath <- "data/eda_final_project"
path <- paste0(p, ath)

setwd(path)
NEI <- readRDS("./summarySCC_PM25.rds")
summary(NEI)
```


